Modal Abstractions for Virtualizing Memory Addresses

### 2024-07-02 examples.tex
+

 \subsubsection{Ensuring Presence of an Entry}
After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
 Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
 In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}. Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.

 There two important aspects on the specification after the call to \textsf{pte\_initialize}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for \textsf{nextpaddr} -- $\ulcorner \Xi !! \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- mapped to a different value (\textsf{Z}) than the identity mapping for the physical address \textsf{entry} is (\textsf{V}). The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line X in Figure \ref{}) because it would create a duplication of the full-ownership of the physical resource in the post-condition would potentially coexists with the one that can be obtained from the side-condition (*) inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi !! \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$. Although, the first aspect recovers the soundness while preserving the existence of the identity mapping for the fresly allocated physical page-table address, unlike  tas the  resource  inside  the  two subtle aspects are important f but quite but quite important for the soundness and 
  \subsubsection{Kernels Utilizing \textsf{P2V}}
  \label{sec:p2v}
 Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we must obtain the virtual address of the next level-page table (Line 46 in \ref{calltopteinitialize}). To do so, as for obtaining any virtual page-table address, we, again, need to open the invariant look up the \textit{identity mapping} for the physical address ($\textsf{entry}\mapsto \textsf{pfn}$).

>>>>

 \subsubsection{Ensuring Presence of an Entry}
After obtaining the identity mapping for \textsf{entry}, we are able to load the \textsf{entry\_val} into \textsf{rdi}, and check the presence bit via $and \;0x1,rdi$.
 Accessing the presence bit and checking the value allow (Lines 7-13 in Figure \ref{fig:calltopteinitialize}) us to exploit the fact in the side condition (*) in such a way that the condition ($\textsf{entry}\mapsto\textsf{present}$ determines whether we take the non-presence path (Lines 12-19 in Figure \ref{fig:calltopteinitialize}) with full ownership ($\textsf{qfrac} = 1$) to set the entry with proper values for the bitfields, e.g., frame showing freshly allocated page and presence bit.
 In case of absence, \textsf{pte\_initialize} is called for a physical-zeroed page (utilizing the page-allocator's \textsf{kalloc} -- currently the only axiom as it is not part of \textsf{pte} (virtual-memory-management) module) to be backing the physical frame address of the entry -- Specification Line 35 in Figure \ref{fig:calltopteinitialize}. Since we are using \textsf{pte\_initialize} for page-table address allocation, it also ensures us that newly allocated physicall address is in the identity mapping map $\Xi$ -- Line X in Figure \ref{fig:calltopteinitialize}.
;;void pte_initialize(pte_t **table) {
;;set up the stack    
;;allocate a full zeroed page for 512 8-byte entries 
;;pte_t *local = kalloc();
callq  81 <pte_initialize+0x11>	7d: R_X86_64_PLT32	kalloc-0x4
mov    rax,-0x10[rbp] 
;;*table = local;
mov    -0x10[rbp],rax
mov    rax,[rdi]
;;return;
;;..
;;clean up the stack
\caption{Allocating a physical page }
\label{pteinitializespec}
 There two important aspects on the specification after the call to \textsf{pte\_initialize} -- Specification in Figure \ref{fig:pteinitializespec}. First, one might have expected to see a physical pointsto for the fresly page-table address  ($\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$) after the call to \textsf{pte\_initialize} but we do not have this physical pointsto resource. Second, we see the identity mapping for \textsf{nextpaddr} -- $\ulcorner \Xi !! \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$ -- mapped to a different value (\textsf{Z}) than the identity mapping for the physical address \textsf{entry} is (\textsf{V}). The first aspect is important for not invalidating the soundness: \textsf{pte\_initialize} extends the identity map with the freshly allocated \textsf{nextpaddr}, and must not reveal the ownership of the physical pointsto in the post-condition (Line X in Figure \ref{}) because it would create a duplication of the full-ownership of the physical resource in the post-condition would potentially coexists with the one that can be obtained from the side-condition (*) inside the extended invariant in Figure \ref{fig:peraspaceinvariant_with_p2v_extension} with the capability token $\ulcorner \Xi !! \textsf{nextpaddr} = \textsf{ Some Z} \urcorner$. Although, the first aspect recovers the soundness while preserving the knowledge of existence of the identity mapping for the fresly allocated physical page-table address, we still need the knowledge that the fresly allocated address references to a \textit{zeroed} page that an explicitly revealed physical pointsto would give -- $\mathsf{nextpaddr} \mapsto_{\mathsf{p}} \mathsf{w64\_0}$. Since a value of an identity mapping inside the invariant is existentially quantified (\textsf{val} in Figure \ref{fig:peraspaceinvariant_with_p2v_extension}), we would loose the knowledge of identity mapping with zeroed value. To carry this knowledge, we distinguish a random value of an identity mapping -- capability representing the existence of the random value (\textsf{V}) inside the invariant for the address \textsf{entry} at Line X in Figure \ref{} -- from the capability representing the zeroed identity-mapping value zeroed (\textsf{Z}). As we do for the address \textsf{nextpaddr}, when we use the capability with value \textsf{Z} for the identity-map lookup ($\Xi !! \textsf{nextpaddr} = \textsf{ Some Z}$), we know that the value existentially quantified inside the invariant is zeroed -- \textsf{val == w64\_0}.
  \subsubsection{Kernels Utilizing \textsf{P2V}}
  \label{sec:p2v}
 Once we make sure that the frame of the page-table entry ($\textsf{entry}\mapsto\textsf{pfn}$)is backed with an allocated physical page, we must obtain the virtual address of the next level-page table (Line 46 in \ref{calltopteinitialize}). To do so, as for obtaining any virtual page-table address, we, again, need to open the invariant look up the \textit{identity mapping} for the physical address ($\textsf{entry}\mapsto \textsf{pfn}$).
